---
title: "hw4 Kutsev Dmitry"
output: html_document
---




```{r, message = TRUE}
#install.packages("tidyverse")
library(tidyverse)

```


```{r, message = TRUE}
tidyverse_packages()
tidyverse_update()

```

```{r, message = TRUE}

mono_socio <- read_csv2("https://raw.githubusercontent.com/LingData2019/LingData2020/master/data/greek-word-order-mono_socio.txt")
mono <- read_csv2("https://raw.githubusercontent.com/LingData2019/LingData2020/master/data/greek-word-order-mono-acceptability-coded-rt.txt")
```


```{r, message = TRUE}
library(ggplot2)
library(ggpubr)

```



Number of participants
```{r, message = TRUE}
nrow(mono_socio['ParticipantID'])

```
Number of participants
```{r, message = TRUE}
mono_socio %>%
  count(ParticipantID!="None")
```
Number of males
```{r, message = TRUE}
mono_socio %>%
  count(Responce=="male")
  
```


Number of females
```{r, message = TRUE}
mono_socio %>%
  count(Responce=="female")
  
```

Which education levels are mentioned in the dataframe?


```{r, message = TRUE}
mono_socio %>%
  filter(QuestionCategory =='education')  %>%
  select(QuestionCategory, Responce) %>%
  distinct(Responce)
 
  
```

How many participants of each education levels are present?

```{r, message = TRUE}
mono_socio %>%
  filter(QuestionCategory =='education')  %>%
  select(QuestionCategory, Responce, ParticipantID) %>%
  count(Responce)
 
  
```

How many left- and right-randed participants are present?

```{r, message = TRUE}
mono_socio %>%
  count(Responce=="left"|Responce=="right")
  
```



```{r, message = TRUE}
mono_socio %>%
  count(Responce=="left"|Responce=="right")
  
```



```{r, message = TRUE}
mono

```
### 1.2 Create a plot that shows the RT distribution in experiment 1 (all participants and conditions taken together). What kind of plot would you choose? Use ggplot() for this problem.

```{r, message = TRUE}
mono %>%
  #group_by(RT, TypeOfQuestion, TypeOfStimuli, ResponseAcceptabilityJudgement, WordOrder) %>%
  #summarise (Freq = n())
  ggplot(aes(x=RT)) + geom_density(alpha=.3)
  #ggplot(summ, aes(x=RT))
  
  
```

### 1.2 Can we say that RT approximately follows normal distribution? Which features of RT distribution contradicts this assumption? (E.g. long left tail, long right tail, outliers, skewness, etc.) 

 - We can see long right tail features.


### 1.3 Normalise data applying the logarithm with base 10 (RTlog = log10(RT)). Use `mutate`.

```{r, message = TRUE}
mono <- mutate(mono, RTlog = log10(RT))
mono
  
```



### 1.4 Create a density plot that shows the RTlog distribution. 

```{r, message = TRUE}
mono %>%
  ggplot( aes(x=RTlog)) +
  geom_density(fill="#69b3a2", color="#e9ecef", alpha=0.8)
  
```


### 1.5
Give a summary of `RTlog` distribution (min, max, mean, median, standard deviation)

```{r, message = TRUE}
summary(mono$RTlog)
```


### 1.6
Filter out outliers. Remove from the table the following observations:
* responses whose RT is below 600 ms (i.e., when a button is pressed too fast, without allowing enough time for actual consideration of the presented stimuli)  
* responses whose RTlog deviates from the mean value of RTlog for more than 3 standard deviations  
* fillers (both acceptable and unacceptable)  
Convert relevant variables to factors and save fitered data as `mono1`.



```{r}
diff <-mean(mono$RTlog)-3*sd(mono$RTlog)
3*sd(mono$RTlog)
mean(mono$RTlog)
diff
mono1 <- mono
mono1 %>% filter(RT > 600, RTlog > diff, TypeOfStimuli != "FillerUnacceptable", 
                 TypeOfStimuli !="FillerAcceptable") -> mono1

mono1 %>% 
  select(ParticipantID, TypeOfStimuli, WordOrder, AcceptabilityJ = 
  ResponseAcceptabilityJudgement, RTlog) %>% 
    mutate(ParticipantID = as.factor(ParticipantID),
      TypeOfStimuli = as.factor(TypeOfStimuli), 
        WordOrder = as.factor(WordOrder), 
        AcceptabilityJ = as.factor(AcceptabilityJ)) -> mono1

```





### 1.7 Calculate the number of observations in `mono1`.

```{r, message = TRUE}
nrow(mono1)
```



### 1.8
Reproduce Figure 1 from the article using `ggplot`.

```{r, message = TRUE}
summ <- mono1 %>%
  group_by(TypeOfStimuli, AcceptabilityJ, WordOrder) %>%
  summarise(Freq = n())
  
  ggplot(summ, aes(x = TypeOfStimuli, y = Freq, fill = AcceptabilityJ)) +
  geom_col(colour = "black", position = "dodge") + facet_wrap(vars(WordOrder))
  
```




### 1.9
Reproduce Figure 2 from the article using ggplot.

```{r, message = TRUE}
summ <- mono1 %>%
  group_by(WordOrder, AcceptabilityJ) %>%
  summarise (Freq = n())
  
  ggplot(summ, aes(x = WordOrder, y = Freq, fill = AcceptabilityJ)) +
  geom_col(colour = "black", position = "dodge")
```


### 1.10
Reproduce Figure 7 from the article using ggplot.

```{r, message = TRUE}
summ <- mono1 %>%
  group_by(TypeOfStimuli, AcceptabilityJ, WordOrder, RTlog) %>%
  summarise (Freq = n())
  
  ggplot(summ, aes(x = AcceptabilityJ, y = RTlog, fill = WordOrder)) +
  geom_violin(colour = "black", position = "dodge") +  facet_wrap(vars(WordOrder)) + 
  geom_boxplot( width=0.1)
```



### 1.11
For the same data, draw a lineplot for group means and standard errors using `ggline()`:


```{r, message = TRUE}
summ <- mono1 %>%
  group_by(TypeOfStimuli, AcceptabilityJ, WordOrder, RTlog) %>%
  summarise (Freq = n())
  
  ggline(summ, x = "AcceptabilityJ", y = "RTlog",  add = c("mean_se"))

```



### 2.1 Summarising
Use `group_by` and `summarise` to find mean logarithm of reaction time for each participant and each word order. Put this dataframe to `mean_rtlog_long` variable. It should be like

```{r, message = TRUE}
mean_mono <- mono1 %>%
  group_by(ParticipantID, WordOrder) %>%
  summarize(mean_rtlog_long = mean(RTlog, na.rm = TRUE))
  #mean_mono
mean_mono
```


### 2.2. Pivoting
Use `pivot_wider` to spread values of `RTlog` in `mean_rtlog_long` into two columns: `Congruent` and `Incongruent`. Put new dataframe in variable `mean_rtlog`. It should look like

```{r, message = TRUE}
mean_mono %>%
  pivot_wider(names_from = WordOrder, values_from = mean_rtlog_long) -> mean_mono_tr
  mean_mono_tr
```


### 2.3. Two-sample t-test
Let us try to apply two-sample t-test to our data. Consider values in columns `Congruent` and `Incongruent` as two independent samples. Our null hypothesis is that these two samples are from populations with equal means. Alternative hypothesis: population mean for incongruate word order is larger (people need more time to ’parse’ it). Use `t.test` function to perform a test. Don't forget to specify `alternative`.

```{r}

mean_mono_diff <- mutate(mean_mono_tr, mean_rtlog = Congruent - Incongruent)
mean_mono_diff 


```

```{r}
t.test(mean_mono_diff$Congruent, mean_mono_diff$Incongruent,var.equal=TRUE, alternative = "greater")

```
Would you reject null hypothesis (under 5% significance level) according to this test?
- No, we would not, p-vlue is more than 0,05.

### 2.4. Paired t-test: manually
To use paired t-test, let us find difference between logarithms of reaction time for each participant. Use `mutate` and add variable `diff` with aforementioned meaning to dataframe `mean_rtlog`. Save result as `mean_rtlog` again. Then compare mean of `diff` with 0 using 1-sample t-test. (Use appropriate alternative.)


```{r}

t.test(mean_mono_diff$mean_rtlog, mu=0)

```

#### 2.5. Paired t-test out of the box
In fact, we can avoid manual calculation of difference and perform paired t-test using `t.test` function with parameter `paired = True`. Apply this function to your data and make sure you get the same result as in 2.4.

```{r}
t.test(mean_mono_diff$Congruent, mean_mono_diff$Incongruent, paired =TRUE)
```



## 3. Difference between conditions
Now we will consider reaction time for Incongruent word ordering only. Let us check are there any statistically significant difference in logarithm of reaction time for different conditions (types of stimuli).



### 3.2 Statistical testing
Use appropriate statistical test to answer the following question: are there any statistically significant difference in logarithm of reaction time for different conditions (types of stimuli)? Choose the test and provide justification for your choice. Provide your code, results and interpretation. What is your answer to the question?

```{r}
res <-aov(summ$RTlog ~ summ$TypeOfStimuli)
res
summary(res)
```
F value is large enough, so there must be statistically significant difference in logarithm of reaction time for different conditions.

### 3.3 Post-hoc analysis: which differences are significant?
If we compare means for several (more than two) groups and reject null hypothesis that corresponding population means are equal to each other, the next natural question is to find all pairs of groups which difference is statistically significant. As we discussed at the lecture, pairwise t-tests cannot be used here without appropriate corrections. Instead, one can use Tukey Honest Significant Differences. It reports adjusted confidence intervals for differences between group means for each pair of groups as well as p-values for null hypothesis ’difference is equal to zero’.

Apply `TukeyHSD` to the result of 3.2 and report which pair of conditions has statistically significant difference between logarithms of reaction time.

```{r}
TukeyHSD(res)
```


















